---
title: "lab_5_rh"
author: "Randi Honeycutt"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objectives

getting counts for different groups
use lubridate to parse dates
find confidence intervervals and do t tests 
heatmap with geom_title()

```{r, include = FALSE}
# attach packages

library(tidyverse)
library(here)
library(janitor)

```

```{r}

#Import lobster data

lobster_abundance <- read_csv(here::here("data","lobster_abundance.csv"), 
                              na = "-99999") %>% 
  clean_names()
```

Use tidyr::uncount() funtion to expand a frequency distribution 

```{r}

lobster_tidy <- lobster_abundance %>% 
  tidyr::uncount(lobster_count)

```

Now each lobster has its own row, this is tidy format. 

Lets look at this in multiple ways:

```{r}

ggplot(data =  lobster_tidy, aes(x = site, y = size_mm)) +
  geom_jitter(aes(color=site), 
              width = 0.2, 
              alpha = 0.3)

ggplot(data =  lobster_tidy, aes(x = site, y = size_mm)) +
  geom_violin(aes(color=site), 
              alpha = 0.3)

ggplot(data =  lobster_tidy, aes(x = size_mm)) +
  geom_histogram(aes(fill = site)) +
  facet_wrap(~site, scales = "free")

ggplot(data =  lobster_tidy, aes(sample = size_mm)) +
  geom_qq(size = 0.5) + 
  facet_wrap(~site, scales = "free")

```

Use 'lubridate' to parse dates and times: 
Add a new column with 'mutate()' that contains my date as an actual date
```{r}

lobster_date <- lobster_tidy %>% 
  mutate(
    date_new = lubridate::mdy(date)
  )

```

Parse 'Date' column to get the different peices (month, year) seperated

```{r}

lobster_parse_date <- lobster_date %>%
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE) , 
    obs_year = lubridate::year(date_new)
  )

```

Count lobsters by different groupings.
I want to count the number of lobsters by year and month 

```{r}

lobster_ym <- lobster_parse_date %>% 
  count(obs_year, obs_month)

lobster_ym

lobster_ysite <- lobster_parse_date %>% 
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>% 
  count(site)

lobster_site

#group_by + summarize + n()
lobster_summary <- lobster_parse_date %>%
  group_by(site) %>% 
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE), 
    sd_size = sd(size_mm, na.rm = TRUE), 
    sample_n = n()
    
  )

```

Confidence intervals and t-tests:

Use the 't.test()' funtion to find confidence intervals and perform t tests 

```{r}

ivee_lobsters <- lobster_tidy %>% 
  filter(site == "IVEE") %>% 
  pull(size_mm)

t.test(ivee_lobsters)
```

two-sample t-test: 
Is there a significant difference in mean lobster lengths b/w naples and mohawk 

```{r}

napl_sample <- lobster_tidy %>% 
  filter(site == "NAPL") %>%
  pull(size_mm)

mohk_sample <- lobster_tidy %>% 
  filter(site == "MOHK") %>% 
  pull(size_mm)

mohk_napl_ttest <- t.test(napl_sample, mohk_sample)
mohk_napl_ttest
  

```

Mean lobster size differed significantly b/w mohawk and naples reefs
(t(`r mohk_napl_ttest$parameter`) = 19.849)

maybe an easier way...

```{r}

lobster_2sample <- lobster_tidy %>% 
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)
```

make a geom_tile heatmap

```{r}

ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) +
  geom_tile(aes(fill = n))


```

